---
title: "Return"
author: "Type your name here!"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
  slidy_presentation: 
    slide_level: 3
  ioslides_presentation:
    slide_level: 3
  beamer_presentation:
    slide_level: 3
---

This is the R markdown (rmd) exercise file that accompanies the lecture notes. It will be your sandbox! Fill in the R code chunks, experiment, and have fun!

# Data Preparation

### Import directly using quantmod


Select any stock other than IBM Download the data using getSymbols() of the quantmod library. Choose a stock that has paid dividend in the past (hint: Google the company name and include "ex-dividend date" in your search). Do not specify "from" and "to", so that quantmod will return the last 10 years of data:
```{r}
library(quantmod)
stock_symbol <- "GS"
stock_data <- getSymbols(stock_symbol, auto.assign = FALSE)
```

# Price Return

### Price Return


See 2.1 of Ang's text.

### Calculate Price Return By Definition


Calculate the daily price return on your stock by definition:
```{r}
Close <- Cl(stock_data)
price.return <- (Close - lag(Close)) / lag(Close)
head(price.return)
```


### Using Diff()


Use Diff() to calculate the daily price return of your stock:
```{r}
price.return <- diff(Close)/lag(Close)
head(price.return)
```


### Using Delt()


Use Delt() to calculate the daily price return of your stock:
```{r}
price.return <- Delt(Close)
head(price.return)
```

### Plot Price Return


Plot the price return of your stock:
```{r}
plot(price.return)
```

# Total Return

### Total Return


Reading: Ch 2.2 of Ang's text.

### Ex-dividend date


Google to find out the most recent ex-dividend date of your stock. (If a stock has never paid dividend, then the exercise here does not apply to the stock.)

Select a time window that include this ex-dividend date and print the stock prices in this time window:
```{r}
cbind(Cl(stock_data), Ad(stock_data))['2018-08-06/2018-08-11']
```

This explains how adjusted close is calculated:

https://help.yahoo.com/kb/SLN28256.html

What is your intuition for the adjustment?

### Calculate Total Return


Calcualte the total return on your stock.
```{r}
Adjusted <- Ad(stock_data)
total.return <- (Adjusted - lag(Adjusted)) / lag(Adjusted)
head(total.return)
```

### Price Return vs. Total Return


Combine the price return and total return into one xts object, and plot the two returns for a subsample that contains the latest ex-dividend date:
```{r}
returns <- cbind(price.return, total.return)
names(returns) <- c('price.return', 'total.return')
plot(returns['2018-08-06/2018-08-15'],
     legend.loc='bottom', major.ticks = "days")
```

### Adjusted Close vs. Unadjusted Close 


Plot the adjusted close and unadjusted close of your stock:
```{r}
plot(cbind(Cl(stock_data), Ad(stock_data)))
```


# Log Return

### Log Return 


Reading: Ch 2.3 of Ang's text.

### Calculate Log Return


Calculate and plot the log return for your stock:
```{r}
log.return <- diff(log(Ad(stock_data)))
plot(log.return)
```

### Log Return and Total Return


Compare log return to total return. How close are they?
They are very close
```{r}
plot(cbind(total.return, log.return))

```

# Cumulative Return

### Cumulative Return


Reading: Ch 2.4 of Ang's text.

### Calculating Gross Return


Calculate the gross return:
```{r}
gross.return <- total.return + 1
head(gross.return)
```

### Calculating Cumulative Return


Calculate the cumulative return:
```{r}
gross.return[1] <- 1
cumulative.return <- cumprod(gross.return)
head(cumulative.return)
```

### Plot the cumulative return


Plot the cumulative return:
```{r}
plot(cumulative.return)
```

How does it compare to the normalized price path?

# Monthly Return

### Monthly Return


Use to.monthly() to convert daily prices to monthly prices:
```{r}
monthly.price <- to.monthly(Adjusted)
head(monthly.price)
```

### Select the end-of-month prices


Select the end-of-month prices:
```{r}
monthly.Close <- monthly.price[,4]
names(monthly.Close) <- paste0(stock_symbol, ".Close")
head(monthly.Close)
```

### Calculate Monthly Return


Calculate monthly returns:
```{r}
monthly.return <- Delt(monthly.Close)
names(monthly.return) <- stock_symbol
head(monthly.return)
```

# Empirical Facts

### Empirical Facts


Remove the NA of your stock returns:
```{r}
r <- na.omit(monthly.return)
head(r)
```


### Monthly Return


Plot the monthly returns:
```{r}
plot(r)
```

## Summary Statistics

### summary()


Find the the first quartile, median, and the third quartile of the stock returns:
```{r}
summary(r)
```
first quartile: -0.052854
median: 0.009394
third quartile: 0.068852
### Range and Interquartile Range


Find the range and interquartile range of the stock returns:
```{r}
range(r)
IQR(r)
```

### Box Plot


Make the box plot of your stock's returns:
```{r}
boxplot(as.vector(r), horizontal = TRUE)
```

Understanding box plot: http://onetipperday.sterding.com/2012/06/about-boxplot.html


### Mean


Calculate the average return of your stock:
```{r}
mean(r)
```

### Variance


Calculate the variance of stock return:
```{r}
var(r)
```

### Standard Deviation


Calculate the standard deviation of stock returns:
```{r}
sd(r)
```

## Are returns normally distributed?

### Histogram


Plot the histogram of the monthly stock return:
```{r}
hist(r)
```


Readings: 

https://www.statmethods.net/graphs/density.html

https://www.r-bloggers.com/basics-of-histograms/

### Change the number of bins


Experiment with the "breaks" option for hist():
```{r}
hist(r, breaks = 20)
```

### Frequency vs Density 


Plot the density instead of frequency:
```{r}
hist(r, breaks = 20, freq = FALSE, main = "Density plot")
```

### Kernel Density


Estimate the kernel density and plot it over the histogram:
```{r}
hist(r, breaks = 20, freq = FALSE, main = "Density Plot")
lines(density(r), col='red', lwd = 2)
```

### Add a Normal Curve


Find help on curve() and dnorm() and use them to plot a normal density on top of the histogram.
```{r}
hist(r, breaks = 20, freq = FALSE, main = "Density Plot")
curve(dnorm(x, mean = mean(r, na.rm = TRUE), sd = sd(r, na.rm = TRUE)), 
      col = "blue", lwd = 2, add = TRUE)
```

What do you observe?
The curve is smoother
### quartile


Use quantile() to return the first quartile, median, and the third quartile of the stock returns:
```{r}
quantile(r)
```

Reading: http://www.r-tutor.com/elementary-statistics/numerical-measures/quartile

### quantile


Find the 5% and 95% quantiles of your stock returns.
```{r}
quantile(r,c(0.05,0.95))
```

Reading: http://www.r-tutor.com/elementary-statistics/numerical-measures/percentile

In spreadsheet, it is accomplished using percentile().

### QQ plot


Make a QQ plot of the stock returns agains a random sample that is normally distributed with the same mean and standard deviation:
```{r}
qqplot(rnorm(length(r), mean(r), sd(r)), as.vector(r))
```


Reading: https://www.dummies.com/programming/r/how-to-use-quantile-plots-to-check-data-normality-in-r/

### QQ plot to check normality


Use qqnorm() to make QQ plot. Are returns normally distributed?
Yes

```{r}
qqnorm(r)
qqline(r, col = "red")
```


### Skewed Distributions


Reading: https://whatis.techtarget.com/definition/skewness

### Measuring Skewness


Calculate the skewness of your stock's returns. You may need to install the e1071 package first.

```{r}
library(e1071)
skewness(r)
```

What is your conclusion?
Mean is less than mode

### Kurtosis


Calculate the kurtosis of your stock's returns. 
```{r}
kurtosis(r)
```

What is your conclusion?
Fat-tailed distribution
Reading: http://www.r-tutor.com/elementary-statistics/numerical-measures/kurtosis


## Return Predictability

### Return Predictability


Make a scatter plot of the stock return against its lag:
```{r}
par(pty="s")
plot(as.vector(r), as.vector(lag(r)))
```
Do past returns predict future returns according to the scatter plot?

### Autocorrelation


Compute the correlation of the return against its lag.
```{r}
cor(r[2:length(r)], lag(r)[2:length(r)])
```

### ACF


Use acf() to plot the autocorrelations:
```{r}
acf(r)
```

What are your findings?

Reading: https://programming-r-pro-bro.blogspot.com/2011/10/predictability-of-stock-returns-using_28.html

### Rolling Volatility


Calcualte the rolling volatility of daily stock returns and intepret:
```{r}
rolling_vol <- rollapply(total.return, width = 30, FUN = sd)
plot(rolling_vol)
```

rollapply() is a very useful function that can be used to calculate rolling sum, rolling mean, etc.

### CBOE Volatility Index (VIX)


Reading: http://www.cboe.com/vix

### Getting VIX Data


The VIX data can be downloaded from: https://www.cboe.com/tradable_products/vix/vix_historical_data/

(If the above link doesn't work, you may need to Google to find out the correct link.)

Download and import the VIX data. (Note: when downloading the data, right click the download link and "save link as" to save to file to the same folder as your rmd file.  Do not open the csv file in Excel to save it to a directory since this will mess up the file format.)
```{r}
vix_data <- read.csv("VIX_History.csv", header = TRUE)
```

### Historical VIX


Plot the historical VIX.  Do you see time-varying volatility?
```{r}
vix_data$DATE <- as.Date(vix_data$DATE, format = "%m/%d/%Y")
vix_xts <- xts(vix_data[,2:5], order.by=vix_data[,1])
plot(vix_xts$CLOSE)
```

