---
title: "Risk"
author: "Type your name here!"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
  slidy_presentation: 
    slide_level: 3
  ioslides_presentation:
    slide_level: 3
  beamer_presentation:
    slide_level: 3
---

This is the R markdown (rmd) exercise file that accompanies the lecture notes. It will be your sandbox! Fill in the R code chunks, experiment, and have fun!

# Value at Risk

### Basic Idea


Simulate your P&L data:
```{r}
PL_simulated <- rnorm(365, mean = 0, sd = 10)
```

Introduction to VaR: https://www.investopedia.com/articles/04/092904.asp

### The Histogram


Estimate the kernel density and plot it over the histogram:
```{r}
hist(PL_simulated, breaks = 29, freq = FALSE)
lines(density(PL_simulated), col = 'red', lwd = 2)
```

### 5% one-day VaR Calculation


Calculate the 5% one-day VaR for the simulated P&L data:
```{r}
(quantile_5_percent <- quantile(PL_simulated, prob = 0.05))
(VaR_5_percent <- -quantile_5_percent)

```

### 1% one-day VaR Calculation


Calculate the 1% one-day VaR for the simulated P&L data:
```{r}
(quantile_1_percent <- quantile(PL_simulated, prob = 0.01))
(VaR_1_percent <- -quantile_1_percent)

```

### VaR Given Realized Returns


Pick your own mutual fund for VaR calculation:

http://eoddata.com/stocklist/USMF.htm

### Mutual Fund Data


Download the mutual fund historical prices from Yahoo Finance:
```{r}
library(quantmod)
symbol <- "BTCRX"
fund.data <- getSymbols(symbol, auto.assign = FALSE, from = "1900-01-01")
```

### Plot the data


Plot the adjusted close:
```{r}
head(fund.data)
plot(fund.data$BTCRX.Adjusted)

```

### Mutual fund returns


Calculate and plot the mutual fund's daily returns:
```{r}
r_p <- Delt(Ad(fund.data)) * 100
r_p <- na.omit(r_p)
names(r_p) <- 'r_p'
plot(r_p)
```

### Summary


Summarize the returns:
```{r}
summary(r_p)
               
```


### Two Methods to Calculate VaR

### Historical VaR


Calculate the historical VaR in percentage terms:
```{r}
(r_p_quantile_5_percent <- quantile(r_p, prob = 0.05))
(historical_VaR_percentage_terms <- - r_p_quantile_5_percent)

```

### From Percentage Terms to Dollar Terms


Choose an initial investment, and convert the percentage VaR to dollar VaR:

```{r}
initial_investment <- 3000
(historical_VaR_dollar_terms <- initial_investment * historical_VaR_percentage_terms / 100) 
```

### The Parametric Approach


Calculate the mean and standard deviation of the returns:
```{r}
(miu <- mean(r_p))
(sig <- sd(r_p))
```


### VaR Calculation: Finding Quantile


Calculate the quantile for a standard normal distribution:
```{r}
(z_quantile <- qnorm(0.05))
```



### VaR Calculation


Calculate the VaR in percentage terms and in dollar terms:
```{r}
r_p_quantile <- miu + sig * z_quantile
(normal_VaR_percentage_terms <- -r_p_quantile)
(normal_VaR_dollar_terms <- initial_investment * normal_VaR_percentage_terms / 100)
```


### VaR at different horizon


Calculate mean and standard deviation at the annual horizon:
```{r}
(miu_annual <- 252 * miu)
(sig_annual <- sqrt(252) * sig)
```


### Annual VaR


Calculate the annual VaR for your investment:
```{r}
(r_p_annual_quantile <- miu_annual + sig_annual * z_quantile)
(normal_VaR_annual_percentage_terms <- -r_p_annual_quantile)
(normal_VaR_annual_dollar_terms <- initial_investment * normal_VaR_annual_percentage_terms / 100)


```


Note that we assume 252 trading days in a year:
https://en.wikipedia.org/wiki/Trading_day

### Historical VaR versus Parametric VaR


Compare the historical VaR and the parametric (normal) VaR:
```{r}
historical_VaR_dollar_terms
normal_VaR_dollar_terms
```
the normal VaR is larger
### Are returns normally distributed?


Plot the histogram and the fitted normal density function to match the mean and standard deviation:

```{r}
hist(r_p, breaks = 500, freq = FALSE)
curve(dnorm(x, mean = mean(r_p), sd = sd(r_p)),
      add=TRUE, col = "darkblue", lwd = 2)
```

### Probability of Observing Extrem Returns


Calculate the tail probability of observing an extreme return (the minimum return), under the assumption that the return is normally distributed:
```{r}
(r_p_min <- min(r_p))
pnorm(r_p_min, mean = miu, sd = sig)
```

# Expected Shortfall

### VaR Shortcomings


VAR versus expected shortfall: https://www.risk.net/risk-magazine/technical-paper/1506669/var-versus-expected-shortfall


### Expected Shortfall (ES)

### ES calculation


Evalvulate the expected short fall, both in percentage terms and in dollar terms:
```{r}
left_tail <- r_p[r_p < r_p_quantile_5_percent]
(ES_percentage_terms <- - mean(left_tail))
(ES_dollar_terms <- initial_investment * ES_percentage_terms/100)
historical_VaR_dollar_terms 

```
